How Kids Code and How We Know:\\An Exploratory Study on the Scratch Repository

### 2016-04-22 Paper/paper.tex
Added blocks per category and type in RQ1

Scratch is a programming language developed to teach children programming by enabling them to create games and interactive animations. The public repository of Scratch programs contains over 12 million projects. Scratch is a \emph{block-based} language: users manipulate blocks to program.

@@

Several studies have shown that block-based languages are powerful as a tool for teaching programming~\cite{meerbaum-salant_learning_2010, b._moskal_evaluating_2005,cooper_teaching_2003}. \todo{Fenia, do you have some work on Research on programming habits}

Recent work has focused on \emph{code smells} within Scratch programs, parts of a program that are not designed in the best possible way. \todo{add dr. scratch papers here} Ar recent controlled experiment found that long scripts and duplication decreases a novice programmer's ability to understand and modify Scratch programs \todo{cite}.

@@

>>>>

Scratch is a programming language developed to teach children programming by enabling them to create games and interactive animations. The public repository of Scratch programs contains over 12 million projects. Scratch is a \emph{block-based} language: users manipulate blocks to program. Block-based languages are visual languages, but also use some successful aspects of text-based languages such as limited text-entry and indentation, and as such are closer to `real', textual programming than other forms of visual programming, like dataflow languages are.

@@

refer to:
Research on the effectiveness of those languages for programming education
Research on programming habits
Our research on smells

@@

We further analyzed the utilization frequency of the different block shapes and categories, as defined in the Scratch documentation. Figures \ref{fig:categories} and \ref{fig:shapes} present the results in terms of number of blocks from the total \nLOC blocks in the dataset projects. The most commonly used blocks are from the Control and Data categories. The Others category includes the blocks related to procedure calls and arguments.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/16Categories}
	\caption{Number of blocks from each category in the analyzed projects}
	\label{fig:categories}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/16Shapes}
	\caption{Number of blocks of each shape in the analyzed projects}
	\label{fig:shapes}
\end{figure}
